---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# flowchart

![travis](https://travis-ci.com/aghaynes/flowchart.svg?token=3En7JHtejF8ywyQBFG5S&branch=master)

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "man/figures/README-"
  )
```

## Goal
The goal of this R package is to make it easier to layout a flowchart. The Gmisc package goes part of the way by allowing one to create boxes with text in and link them together, but you still have define the locations of boxes yourself. This is where flowchart helps. It automatically places the boxes in the figure space by making suitable guesses based on the data supplied it. 


## Installation/Loading
This package is currently only available on github.

```{r inst, eval=FALSE}
# install.packages("devtools")
# less developmental version
devtools::install_github("CTU-bern/flowchart")
# main developmental version
devtools::install_github("aghaynes/flowchart")

```
The package is loaded into R as usual
```{r, warning=FALSE}
library(flowchart)
```

## Usage

`flowchart` makes it's guesses based on data provided in the form of a dataframe. Each row in the dataframe represents a box in the flowchart. `flowchart` expects, as a minimum, columns named `level`, `group`, `text` and `incexc`. These four variables represent the verticle position (from the top), the horizontal position (from the left, or `NA` for central), the text that should be in the box, and whether the row is for inclusion (`"inc"`) or exclusion (`"exc"`).

```{r}
dat <- data.frame(level = c(1, 2, 3, 4, 4, 5, 5, 6, 6),
                  group = c(NA, NA, NA, 1, 2, 1, 2, 1, 2),
                  text  = c("Assessed for eligibility", "Not eligible\nfor inclusion", 
                            "Randomized", "Group1", "Group2", "Excl 1", "Excl 2", "Analysis1", 
                            "Analysis2"),
                  incexc = c("inc", "exc", "inc", "inc", "inc", "exc", "exc", 
                             "inc", "inc"))
dat
flowchart(dat)
```


In the above example, the exclusion boxes are a little too far to the left. This is easily rectified.
```{r}
flowchart(dat, group_exc_shift = 0.5)
```

Note that one box has two lines of text. This is accomplished with the newline operator (`\n`).

### Changing locations of boxes
`flowchart` provides the option to return a dataframe of values that can be easily modified to adjust the position of boxes (note that the figure has been suppressed in this document, it is always produced by `flowchart`).
```{r, fig.show='hide'}
tab <- flowchart(dat, tab = TRUE)
tab
```

The `x` and `y` columns are in proportions of the device space and represent the middle of the box. The `width` column is the width of the boxes, again in proportions. Thus, to decrease the widths of the "Excl" boxes and move the right one further to the right, we can edit those values in `tab` and use tab in `flowchart`:
```{r}
tab$x[7] <- 0.8
tab$width[6:7] <- 0.1
flowchart(tab)
```


### Special fonts and characters
`flowchart` can also use expressions to create bold or italic fonts, or to include Greek letters. 
```{r}
tab$text[4] <- "expression(paste(bold('Group'), '1'))"
tab$text[5] <- "expression(paste(italic('Group'), '2'))"
flowchart(tab, group_exc_shift = 0.5)
```


